\mytitle

### 2020-04-30 paper.tex
Edits to end of Sect 4.1

WFL-QCOP can be solved using an exact optimization solver, e.g., Gurobi~\cite{gurobi}. 
Similar, linear formulations have been previously proposed and solved using exact solvers~\cite{Zhang2014,donovan2005wind}.
\todo{JCB: We need to formally present the ILP model in the experiment section. We can then discuss the fact that the model is identical - because the linearization gives and exact representation. Should put a forward reference to that model here.}
%\todo{Are these linear formulations ``equivalent'' as you claimed or approximations? If they are mathematically equivalent, why do we think a quadratic model will be any better computationally? Arik: The two formulations are equivalent to WFL-QCOP. However, we compare against the ILP model experimentally and it does worse than WFL-QCOP. Hopefully, Jason's part in Section 3.5.1. based on Wen-Yang's thesis will explain why it does better. Originally, I used the quadratic formulation because it seemed more natural. Semantically, the linear model and the quadratic model are the equivalent imho.}

@@

To test and benchmark our models, we used 12 standard benchmark WFLO instances
from the literature~\cite{MOSETTI1994105,Zhang2014,grady2005placement}. 
The 12 benchmarks result from varying 2 wind settings (1 and 36 wind directions),
2 grid sizes ($10\times10$ with cell size $200$ meters and $20\times20$ with cell size $100$ meters),
and 3 variations for number of turbines ($m \in \{20, 30 or 40\}$). 
For the $36$ wind directions setting, the velocities came from the probability distribution presented in Figure~\ref{fig:prob_wind}. We denote $\mathcal{W}$ the set of 12 instances. Note that instances with $36$ wind regimes are considered computationally harder~\cite{Zhang2014}.

@@

We compared our two models, namely WFL-QCOP and WFL-QUBO, against two state of the art exact optimization methods:
the integer linear programming (ILP-LS) approach from~\cite{donovan2005wind}, and the 
model that uses a quadratic approximation (QP-SS)~\cite{turner2014new}. The CP approach presented in~\cite{Zhang2014} was
reported to perform worse than the rest and was left out of the experiments.
We used state of the art software, namely Gurobi~\cite{gurobi}, to solve WFL-QCOP, ILP-SS and QP-SS. 
For consistency, we shall refer to WFL-QCOP problem as QP-LS as it is a quadratic program that uses the LS objective.
In addition, we 
applied both Fujitsu's digital annealer and Gurobi to solve WFL-QUBO: we shall refer to the former as DA-LS, and to the latter
as QUBO-LS.

@@

For exact methods (QP-SS, ILP-LS, QP-LS and QUBO-LS),
we ran Gurobi with a time limit of 3600 seconds. We collected all feasible solutions
until that point in time. The best solution in terms of energy is the solution that attained the highest objective
values (SS in QP-SS and LS in the rest of the methods).
For the digital annealer, we used a single configuration that was found to work best across the different instances. The initial temperature is 500000, the temperature mode is 0, the  interval for updating is 1, the offset increase rate is 10, the number of runs is 20, and the temperature decay rate is computed dynamically. \todo{Eldan and Arik: Do I need to write the formula for computing temperature decay rate? -- Jason} The DA ran until convergence of the underlying simulated annealing algorithm and the runtime 
of the procedure was recorded. \todo{@Chris: do we want to share the code that ran the experiments in batches? perhaps it should be separated from the problem construction and the two need to be shared separately}

The exact methods were all solved using Gurobi's quadratic programming solver on a Ubuntu PC with Inte(R) Core(TM) i7-9700 8 core CPU @3.00 GHz with 12 MB cache and 8 GB RAM. Furthermore, QUBO-LS was solved on DA version 1 \todo{Eldan: any additional info that we need to add here?}.

@@

We have measured the total expected energy of a solution using the SS expression (see Eq.~(\ref{eq:ss}))\footnote{Note that the total mean energy is presented in kW.}.
The only method that optimizes on SS is QP-SS. Hence, for the other methods, we converted solutions
LS-based solution into the SS expression using Eq.~(\ref{eq:ss}) as proposed in~\cite{Zhang2014}. Specifically, let $x = (x_1, \ldots, x_k), k=n^2$ be a turbine placement solution obtained by an LS-based method.
Then, the SS based energy expression can be written as, 
$$\sum_{i\in \mathcal{N}}\sum_{d \in \mathcal{D}} \frac{1}{3} x_i \Bigg(u_{id,\infty} \Bigg[1-\sqrt{\sum_{j\in \mathcal{U}_{id}}x_j\bigg( 1-\frac{u_{ijd}}{u_{id,\infty}} \bigg) } \Bigg]    \Bigg)^3.$$

Next, we tested the performance of our approach as function of solution time. To this end, we 
compared the average performance of all WFLO methods on all 12 WFLO instances. Since energy levels are not directly comparable between the different instances (due to varying number of turbines and wind farm specs) we used the \emph{mean relative error} (MRE) measure. The MRE is defined with respect to the best solution obtained per instance and can be computed over time. Then, the
time-varying MRE can be averaged across the 12 scenarios and presented over time. In what follows, we demonstrate the construction of the MRE measure.

Recall that $\mathcal{W}$ is the set of our 12 standard WFLO instances. 
Moreover, we let $b_w$ be the best energy solution (in terms of SS) obtained for an instance $w \in \mathcal{W}$ 
across the different methods. For example, the best solution (within 3600 seconds) for the second instance $w_2$ that aims at placing $30$ turbines on a $10\times10$ grid with a single wind regime is attained by QUBO-LS and the SS energy expression is $15742$ (kW); thus, we set $b_w = 15742$. Then, for every point in time $t \in \{0,\ldots, 3600\}$ (measured in seconds), we can compute the relative error of a given approach per instance $w$. Let $\mathcal{A}$ be the set of solution approaches, $\mathcal{A}= \{QUBO-LS, QP-LS, DA-LS, ILP-LS, QP-SS\}$. Further, denote by $b_{w,t,a}$ denote the best solution attained before time $t$ of approach $a$ for instance $w$. Then, the mean relative error at time $t$ for approach $a$ on instance $w$ is given by \begin{equation}MRE(w,t,a) = \frac{b_w - b_{w,t,a}}{b_w}.\end{equation} Note that the expression is always positive, since we are solving a maximization problem.

We can now average over the different instances at time $t$. The average performance of approach $a$ at time $t$, $MRE(t,a)$ can be computed as, \begin{equation} MRE(t,a) = \frac{1}{|\mathcal{W}|}\sum_{ w\in\mathcal{W}} MRE(w,t,a).\end{equation}
We shall present $MRE(t,a)$ for each of the approaches to demonstrate the speed of convergence of the proposed LS-based methods.

>>>>

WFL-QCOP can be solved using an exact optimization solver, e.g., Gurobi~\cite{gurobi}. 
Similar, linear formulations have been previously proposed and solved using exact solvers~\cite{Zhang2014,donovan2005wind}.
%\todo{JCB: We need to formally present the ILP model in the experiment section. We can then discuss the fact that the model is identical - because the linearization gives and exact representation. Should put a forward reference to that model here.}
%\todo{Are these linear formulations ``equivalent'' as you claimed or approximations? If they are mathematically equivalent, why do we think a quadratic model will be any better computationally? Arik: The two formulations are equivalent to WFL-QCOP. However, we compare against the ILP model experimentally and it does worse than WFL-QCOP. Hopefully, Jason's part in Section 3.5.1. based on Wen-Yang's thesis will explain why it does better. Originally, I used the quadratic formulation because it seemed more natural. Semantically, the linear model and the quadratic model are the equivalent imho.}

@@

To test and benchmark our models, we used 12 standard benchmark WFLO instances
from the literature~\cite{MOSETTI1994105,Zhang2014,grady2005placement} that we denote with the set $\mathcal{W}$. 
The 12 benchmarks result from varying 2 wind settings (1 and 36 wind directions),
2 grid sizes ($10\times10$ with cell size $200$ meters and $20\times20$ with cell size $100$ meters),
and 3 variations for number of turbines ($m \in \{20, 30 or 40\}$). 
For the $36$ wind directions setting, the velocities came from the probability distribution presented in Figure~\ref{fig:prob_wind}.
%We denote $\mathcal{W}$ the set of 12 instances.
Instances with $36$ wind regimes are considered computationally more challenging~\cite{Zhang2014}.

@@

We compared our two models, WFL-QCOP and WFL-QUBO, against two state-of-the-art exact optimization methods:
the integer linear programming (ILP-LS) approach~\cite{donovan2005wind,Zhang2014} and the 
model that uses a quadratic approximation (QP-SS)~\cite{turner2014new}. 
The ILP model is a standard, exact linearization of the WFL-QCOP model.
%% given in the following set of equations, which we denote WFL-ILP, \begin{eqnarray} \label{ILP}
%% &\max_{x_i}^{}& \sum_{i \in \mathcal{N}}^{} \sum_{d \in \mathcal{D}}^{} p_d ( \frac{1}{3} \ u_{id, \infty}^3 x_i  - \sum_{j \in \mathcal{U}_{id}}^{} \frac{1}{3}(u_{id, \infty} ^3 - u_{ijd}^3)y_{i,j})   \\
%% &s.t.:& \nonumber\\
%% &\mbox{       }& \sum_{i \in \mathcal{N}}^{} x_i = m,\label{Cardinality2}\\
%% &\mbox{       }& x_i + x_j \leq 1,   \forall (i,j) \in \mathcal{E}, \label{Grid2}\\
%% &\mbox{       }& x_i + x_j -1 \leq y_{i,j},  \forall i,j \in \mathcal{N}, \label{Linearization}\\
%% &\mbox{       }& y_{i,j} \geq 0,     \forall i,j \in \mathcal{N}, \label{Yij}\\ 
%% &\mbox{       }& x_i \in \{0,1\},     \forall i \in \mathcal{N}. \label{Vars2} 
%% \end{eqnarray} The variables $y_{i,j}$ are continuous and can take only non negative values (Eq.(\ref{Yij})). However, due to Eq.~(\ref{Yij}) their values must
%% be either zero or one. If a turbine is placed in both $i$ and $j$ ($x_i = x_j = 1$), then $y_{i,j}$ will be $1$ and a wake reduction 
%% will take place in the objective function. If a turbine is placed in either $x_i$ or $x_j$ (but not in both), $y_{i,j}$ will
%% take the value of $0$ and no wake reduction will occur. The WFL-ILP is equivalent to WFL-QCOP: the interaction
%% variables $y_{i,j}$ of the ILP model are replaced by multiplicative expressions $x_{i}x_{j}$, which makes Constraint (\ref{Yij}) redundant. 
However, the ILP is solved using linear optimization approaches, while WFL-QCOP is solved using quadratic programming methods as mentioned in Section~\ref{quadratic}. Hence
we can expect a difference in the performance of optimization software when solving the two problems.

The existing constraint programming approach was
reported to perform worse than the ILP model~\cite{Zhang2014} and so was omitted from our the experiments.
We used state-of-the-art commercial software, Gurobi~\cite{gurobi}, to solve WFL-QCOP, ILP-SS and QP-SS. 
For consistency, we shall refer to WFL-QCOP problem as QP-LS as it is a quadratic program that uses the LS objective.
In addition, we 
used both Fujitsu's Digital Annealer and Gurobi to solve WFL-QUBO: we shall refer to the former as DA-LS, and to the latter
as QUBO-LS.

@@

For exact methods (QP-SS, ILP-LS, QP-LS and QUBO-LS),
we ran Gurobi with a time limit of 3600 seconds. We collected all feasible solutions
found until that point in time.
%The best solution in terms of energy is the solution that attained the highest objective
%values (SS in QP-SS and LS in the rest of the methods).
For the DA, we used a single configuration that was found to work best across the different instances. The initial temperature is 500000, the temperature mode is 0, the  interval for updating is 1, the offset increase rate is 10, the number of runs is 20, and the temperature decay rate is computed dynamically. \todo{Eldan and Arik: Do I need to write the formula for computing temperature decay rate? -- Jason} \todo{JCB: Yes, if it is not too complicated. You should mention if this is the standard thing that DA does or something you came up with.} The DA ran until convergence of the underlying simulated annealing algorithm and the runtime 
of the procedure was recorded. \todo{@Chris: do we want to share the code that ran the experiments in batches? perhaps it should be separated from the problem construction and the two need to be shared separately.} \todo{JCB: As I requested last fall, we should have separate problem instance files from the code and make at least the problem instances available.}

The exact methods were all solved using Gurobi's quadratic programming solver on a Ubuntu PC with Inte(R) Core(TM) i7-9700 8 core CPU @3.00 GHz with 12 MB cache and 8 GB RAM. Furthermore, QUBO-LS was solved on DA version 1. \todo{Eldan: any additional info that we need to add here?}

@@

We measured the total expected energy of a solution using the SS expression (see Eq.~(\ref{eq:ss}))\footnote{The total mean energy is presented in kW.}.
The only method that directly optimizes SS is QP-SS. Hence, for the other methods, we converted solutions
LS-based solution into the SS expression using Eq.~(\ref{eq:ss}) as proposed in Zhang et al.\ \cite{Zhang2014}. Specifically, let $x = (x_1, \ldots, x_k), k=n^2$ be a turbine placement solution obtained by an LS-based method.
Then, the SS based energy expression can be written as, 
$$\sum_{i\in \mathcal{N}}\sum_{d \in \mathcal{D}} \frac{1}{3} x_i \Bigg(u_{id,\infty} \Bigg[1-\sqrt{\sum_{j\in \mathcal{U}_{id}}x_j\bigg( 1-\frac{u_{ijd}}{u_{id,\infty}} \bigg) } \Bigg]    \Bigg)^3.$$

Next, we tested the performance of our approach as function of solution time. We 
compared the average performance of all WFLO methods on all 12 WFLO instances. Since energy levels are not directly comparable between the different instances (due to varying number of turbines and wind farm specifications) we used the \emph{mean relative error} (MRE) measure. The MRE is defined with respect to the best solution obtained per instance and can be computed over time. Then, the
time-varying MRE can be averaged across the 12 scenarios and presented over time. In what follows, we demonstrate the construction of the MRE measure.

Recall that $\mathcal{W}$ is the set of our 12 standard WFLO instances. 
Let $b_w$ be the best energy solution (in terms of SS) obtained for an instance $w \in \mathcal{W}$ 
across the different methods.
%For example, the best solution (within 3600 seconds) for the second instance $w_2$ that aims at placing $30$ turbines on a $10\times10$ grid with a single wind regime is attained by QUBO-LS and the SS energy expression is $15742$ (kW); thus, we set $b_w = 15742$.
Then, for every point in time $t \in \{0,\ldots, 3600\}$ (measured in seconds), we can compute the relative error of a given approach per instance $w$. Let $\mathcal{A}$ be the set of solution approaches, $\mathcal{A}= \{QUBO-LS, QP-LS, DA-LS, ILP-LS, QP-SS\}$. Further, denote by $b_{w,t,a}$ the best solution attained before time $t$ of approach $a$ for instance $w$. Then, the mean relative error at time $t$ for approach $a$ on instance $w$ is given by \begin{equation}MRE(w,t,a) = \frac{b_w - b_{w,t,a}}{b_w}.\end{equation} Note that the expression is always positive, since we are solving a maximization problem.

We can now average over the different instances at time $t$. The average performance of approach $a$ at time $t$, $MRE(t,a)$ can be computed as, \begin{equation} MRE(t,a) = \frac{1}{|\mathcal{W}|}\sum_{ w\in\mathcal{W}} MRE(w,t,a).\end{equation}
We present $MRE(t,a)$ for each of the approaches to demonstrate the speed of convergence of the proposed LS-based methods.
